---
title: "PA1_template"
author: "Soumava Dey"
date: "July 30, 2018"
output: html_document
---
## Loading and preprocessing the data

Load the data using read.csv() function

```{r setup, echo=TRUE}

setwd("C:/Users/soudey/Documents/Data Science/RepData_courseassignment1")

activity_dataset <- read.csv("activity.csv")

head(activity_dataset)

```



## What is mean total number of steps taken per day?

Calculate the total number of steps taken per day

Make a histogram of the total number of steps taken each day

```{r activity, echo=FALSE}
library(ggplot2)

activity_steps_day <- tapply(activity_dataset$steps,activity_dataset$date,sum,na.rm= TRUE)

qplot(activity_steps_day,xlab="Steps per day", ylab="Frequency",col="Red",binwidth=500)

```

Calculate and report the mean and median of the total number of steps taken per day

```{r mean_median, echo=TRUE}
activity_steps_day_mean <- mean(activity_steps_day)
activity_steps_day_median <- median(activity_steps_day) 

head(activity_steps_day_mean)
head(activity_steps_day_median)

```

## What is the average daily activity pattern?

Make a time series plot

```{r time_series, echo=FALSE}

time_series <- aggregate(x = list(meansteps=activity_dataset$steps), by=list(meaninterval = activity_dataset$interval), FUN = mean, na.rm = TRUE)

ggplot(time_series, aes(x = meaninterval, y = meansteps)) +
  geom_line() +
  xlab("5-minute interval") +
  ylab("No. Of steps")

```



Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r max_steps, echo=FALSE}


max_interval<- which.max(as.numeric(time_series$meansteps))

head(max_interval)

```

##Imputing missing values 

Calculate the total number of missing values

```{r missung_values, echo=FALSE}

activity_NA <- sum(is.na(activity_dataset))

head(activity_NA)

```

Devise a strategy to filling up all the missing values

```{r missung_values_impute, echo=FALSE}

library(Hmisc)

activity_impute <- activity_dataset

activity_impute$steps <- impute(activity_dataset$steps, fun=mean)

```

Make a histogram of the total number of steps taken each day (imputed data)

```{r activity_imputed, echo=FALSE}
activity_steps_day_imputed <- tapply(activity_impute$steps,activity_dataset$date,sum)

qplot(activity_steps_day_imputed,xlab="Steps per day(imputed)",ylab="Frequency",col="Red",binwidth=500)

```


Mean and median of steps based on imputed data

```{r imputed mean_median, echo=TRUE}
activity_steps_day_mean_imputed <- mean(activity_steps_day_imputed)
activity_steps_day_median_imputed <- median(activity_steps_day_imputed) 

head(activity_steps_day_mean_imputed)
head(activity_steps_day_median_imputed)

```

##Are there differences in activity patterns between weekdays and weekends?

```{r weekday_weekend_pattern, echo=FALSE}
activity_impute$date <- as.Date(activity_impute$date )
#create a vector of weekdays
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
#Use `%in%` and `weekdays` to create a logical vector
#convert to `factor` and specify the `levels/labels`
activity_impute$wDay <- factor((weekdays(activity_impute$date) %in% weekdays1), 
         levels=c(FALSE, TRUE), labels=c('weekend', 'weekday')) 
#Or
activity_impute$wDay <- c('weekend', 'weekday')[(weekdays(activity_impute$date) %in% weekdays1)+1L]

averagedActivityDataImputed <- aggregate(steps ~ interval + wDay, data=activity_impute, mean)

ggplot(averagedActivityDataImputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(wDay ~.) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")


```





